---

title: "Estimating a Bernstein Copula from Multivariate Normal Data"
output: rmarkdown::html\_vignette
vignette: >
%\VignetteIndexEntry{BernsteinCopula: Estimation from Multivariate Normal}
%\VignetteEngine{knitr::rmarkdown}
\usepackage\[utf8]{inputenc}
----------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6)
library(BernsteinCopula)
library(MASS)
library(ggplot2)
```

## Introduction

In this vignette, we demonstrate how to estimate a Bernstein copula from samples drawn from a bivariate normal distribution. We jointly estimate the copula structure and marginal distributions via the EM algorithm.

## Simulating Data from a Bivariate Normal

```{r simulate-data}
set.seed(123)
Sigma <- matrix(c(1, 0.7, 0.7, 1), 2, 2)
data <- MASS::mvrnorm(n = 200, mu = c(0, 0), Sigma = Sigma)
x <- exp(data[, 1])
y <- exp(data[, 2])
```

## Visualizing the Data

```{r plot-data}
df <- data.frame(x = x, y = y)
ggplot(df, aes(x = x, y = y)) +
  geom_point(alpha = 0.6) +
  labs(title = "Bivariate Normal Data", x = "X", y = "Y") +
  theme_minimal()
```

## Initializing Marginal Models

We use `normaldist()` to create EM-aware marginal distributions for both `x` and `y`.

```{r marginal-models}
Fx <- normaldist(mean = mean(x), sd = sd(x))
Gy <- normaldist(mean = mean(y), sd = sd(y))
```

## EM Estimation with Marginals and Copula

We estimate the Bernstein copula and marginal parameters jointly using `emloop_with_F()`.

```{r emloop-joint}
m <- 4
n <- 4
R0 <- matrix(1 / (m * n), m, n)

result <- emloop_with_F(
  x = x,
  y = y,
  R_init = R0,
  F = Fx,
  G = Gy
)

print("Estimated parameters only from marginals:")
print("Fx params:")
print(Fx$params)
print("Gy params:")
print(Gy$params)

print("Estimated parameters with joint copula:")
print("Fx params:")
print(result$params1)
print("Gy params:")
print(result$params2)

Rhat <- result$R
Fx$params <- result$params1
Gy$params <- result$params2
```

## Visualizing Updated Marginals

```{r updated-marginals}
xgrid <- seq(min(x) - 2, max(x) + 2, length.out = 200)

df_fx <- data.frame(x = xgrid, density = Fx$pdf(xgrid))
ggplot(df_fx, aes(x = x, y = density)) +
  geom_line(color = "steelblue") +
  labs(title = "Updated Marginal of X", x = "X", y = "Density") +
  theme_minimal()

df_gy <- data.frame(y = xgrid, density = Gy$pdf(xgrid))
ggplot(df_gy, aes(x = y, y = density)) +
  geom_line(color = "darkgreen") +
  labs(title = "Updated Marginal of Y", x = "Y", y = "Density") +
  theme_minimal()
```

## Visualizing the Estimated Copula Density

```{r plot-copula-density}
xg <- seq(min(x), max(x), length.out = 50)
yg <- seq(min(y), max(y), length.out = 50)
Z <- bernstein_joint_pdf_grid(xg, yg, Rhat, Fx, Gy)

df_data <- data.frame(x = x, y = y)

plot_contour(xg, yg, Z,
             title = "Estimated Joint Density (filled)",
             xlabel = "x", ylabel = "y")

plot_contour(xg, yg, Z, filled = FALSE,
             title = "Estimated Joint Density with Data",
             xlabel = "x", ylabel = "y",
             points = df_data)
```

## Conclusion

We demonstrated how to estimate a Bernstein copula jointly with marginal distributions using the `emloop_with_F()` function. This approach captures both dependence and marginal behavior from data using the EM algorithm.
