---
title: "Estimating a Bernstein Copula from Multivariate Normal Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BernsteinCopula: Estimation from Multivariate Normal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6)
library(BernsteinCopula)
library(MASS)
library(ggplot2)
```

## Introduction

In this vignette, we demonstrate how to estimate a Bernstein copula from samples drawn from a bivariate normal distribution.

## Simulating Data from a Bivariate Normal

We use the `MASS::mvrnorm` function to generate samples:

```{r simulate-data}
set.seed(123)
Sigma <- matrix(c(1, 0.7, 0.7, 1), 2, 2)
data <- MASS::mvrnorm(n = 200, mu = c(0, 0), Sigma = Sigma)
x <- data[,1]
y <- data[,2]
```

## Visualizing the Data

We can visualize the simulated data using a scatter plot:
```{r plot-data}
df <- data.frame(x = x, y = y)
ggplot(df, aes(x = x, y = y)) +
  geom_point(alpha = 0.6) +
  labs(title = "Bivariate Normal Data", x = "X", y = "Y") +
  theme_minimal()
```

## Estimating Marginal Distributions
We can estimate the marginal distributions of `x` and `y` using the normal distribution:
```{r marginal-distributions}
Fx <- normaldist(mean = mean(x), sd = sd(x))
Gy <- normaldist(mean = mean(y), sd = sd(y))

xgrid <- seq(min(x) - 2, max(x) + 2, length.out = 200)

df_fx <- data.frame(x = xgrid, density = Fx$pdf(xgrid))
ggplot(df_fx, aes(x = x, y = density)) +
  geom_line(color = "steelblue") +
  labs(title = "Marginal Distribution of X", x = "X", y = "Density") +
  theme_minimal()

df_gy <- data.frame(y = xgrid, density = Gy$pdf(xgrid))
ggplot(df_gy, aes(x = y, y = density)) +
  geom_line(color = "darkgreen") +
  labs(title = "Marginal Distribution of Y", x = "Y", y = "Density") +
  theme_minimal()
```

## Transforming to Uniform Margins

To estimate the copula, we first transform the data to uniform margins using the cumulative distribution function (CDF) of the normal distribution:

```{r pseudo-observations}
u <- Fx$cdf(x)
v <- Gy$cdf(y)
```

## EM Estimation of the Bernstein Copula

We initialize a uniform matrix for the copula weights:

```{r em-estimation}
m <- 4
n <- 4
R0 <- matrix(1/(m * n), m, n)
result <- emloop(u, v, R_init = R0)
result$R
```

## Visualizing the Copula Density

We now transform the original data to copula density estimates:

```{r plot-density}
xg <- seq(min(x), max(x), length.out = 50)
yg <- seq(min(y), max(y), length.out = 50)
Z <- bernstein_joint_pdf_grid(xg, yg, result$R, Fx, Gy)

df_data <- data.frame(x = x, y = y)
plot_contour(xg, yg, Z,
             title = "Estimated Joint Density with Data",
             xlabel = "x", ylabel = "y")

plot_contour(xg, yg, Z, filled=FALSE,
             title = "Estimated Joint Density with Data",
             xlabel = "x", ylabel = "y",
             points = df_data)
```

## Conclusion

We have shown how to estimate a Bernstein copula from bivariate normal data using the EM algorithm. The resulting copula captures the dependence structure and can be visualized with estimated joint density plots.
